MSc Thesis

### 2023-01-27 manuscript/main.tex
eliminated passive voice like a boss

Silicon abounds in the earth’s crust, with silicon dioxide (SiO\textsubscript{2}) comprising about 60\% of the crust by mass. Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant. Silicon can be deposited throughout the whole plant, forming amorphous masses of silica, with highly variable geometries (\cite{piperno_phytoliths_2006}). Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation. When supplemented with silicon, plants are generally more resistant to stress. Silicon supplementation shows efficacy in relieving the negative effects of such abiotic stresses as: soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}). Silicon is also effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}). Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}). The recently put-forward apoplastic barrier hypothesis suggests that the various benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition (\cite{coskun_controversies_2019}). Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector proteins, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology.The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wear down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}). Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.

@@

To test the prevalence of rapid silicon accumulation in canadian cereal crops, I selected three cultivars for each of oats, bread wheat, triticale, and barley. Cultivars were selected on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, I germinated seeds in germination trays filled with moist sand. After four days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid. Though potting mix and fresh water contain some amount of plant available silicon, the silicic acid was added to ensure that there would be no silicon limitation to the plants. I randomized the location of each pot within the growing space. The pots were bottom watered on flood tables with nutrient solution. I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. Simulated herbivory was achieved by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while true herbivory was provided by crickets housed in water-pik tubes. Prior to introduction to the the plants, I acclimated crickets by feeding them on the same species used in this trial. Immediately preceding cricket application, I placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.

@@

24 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e. no cricket exposure) leaves. Leaf tissue was flash frozen in liquid nitrogen, freeze dried, and stored in a -80ºC freezer. To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube. I then flash froze the tubes in liquid nitrogen and ground the leaf tissue in a tissuelyser bead mill for 30 seconds at 60 Hz.

@@

To measure the silicon content of the leaf tissue, I followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}). I pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet. I then placed the pellet in the XRF hood, and used a 30 second scan time to quantify silicon. The pellet die and the analyzer were wiped clean after each use to minimize contamination between samples.

>>>>

Silicon abounds in the earth’s crust, with various silicates, such as silicon dioxide (SiO\textsubscript{2}), comprising about 60\% of the crust by mass. Nearly all terrestrial plants grow in soils containing silicon, and thus absorb nominal amounts through passive transport as water is absorbed into the plant. Silicates are found in a variety of forms, and vary in their plant availability. Crystalline forms, such as quartz, are highly resistant to weathering, and are poor sources of plant-available silicon, while amorphous forms of SiO\textsubscript{2} are more available. As soils age, an increasing share of the plant-available silicon is derived from biogenic silicates, such as diatom testes or plant phytoliths. Plant phytoliths are amorphous masses of silica, generally similar to opal, that are found throughout the plant body and have highly variable geometries (\cite{piperno_phytoliths_2006}). Phytolith morphologies are generally conserved within taxonomic groups, allowing their use as a tool for paleobotanical investigations. The highly variable morphology among clades, with the conservation of morphology within clades suggests active selection upon the structure of phytoliths. The morphology of phytoliths also varies between organs in the plant. Stem phytoliths tend to be more elongate, with putative structural function, while leaf phytoliths are typically more orbicular, and are likely an adaptation to counter herbivory.

Though widespread benefits of silicon are readily observed in many vascular plant families, the exact mechanism through which silicon acts is still poorly understood (\cite{coskun_controversies_2019}), in part due to the seemingly disparate traits that are promoted under silicon supplementation. When supplemented with silicon, plants are generally more resistant to stress. Silicon supplementation shows efficacy in relieving the negative effects of such abiotic stresses as: soil salinity, soil metal toxicity, cold and heat stress, UV stress, water deficits, and phosphorus deficiencies (\cite{cooke_consistent_2016}). Silicon is also effective at limiting the growth and damage of insect and fungal pests (\cite{fauteux_silicon_2005,massey_herbivore_2007}). Silicon-supplied plants put under stress show transcriptome profiles similar to unstressed plants (\cite{coskun_controversies_2019}). The recently put-forward apoplastic barrier hypothesis suggests that the various benefits gained under silicon supplementation derive from the toughening and sealing action of silicon deposition (\cite{coskun_controversies_2019}). Silicon deposits reduce water loss and radiation/temperature damage, and also limit the spread of effector proteins, dampening the effects of fungal pathogen and herbivore excretions designed to interfere with plant defensive physiology.The toughness of the depositions also serves a more direct mechanical role, as the hardened granules of silicon interrupt the chewing motions of herbivores, wear down mandibles and teeth (\cite{stromberg_functions_2016,waterman_short-term_2021-1}), and reduce the digestive efficiencies of herbivores (\cite{johnson_silicon_2021}). Continuing to untangle the various mechanisms through which silicon delivers beneficial effects to plants is key to fully realizing the potential of silicon in sustainable agriculture.

@@

To test the prevalence of rapid silicon accumulation in canadian cereal crops, I selected three cultivars for each of oats, bread wheat, triticale, and barley. I selected cultivars on the basis of minimizing shared pedigree, and no cultivars shared more than one common ancestor within the last two crossing generations. At the start of the experiment, I germinated seeds in germination trays filled with moist sand. After four days, I transplanted germinated seedlings into 10cm pots filled with SunGro potting mix amended with [amount] of silicic acid. Though potting mix and fresh water contain some amount of plant available silicon, I added the silicic acid to ensure that there would be no silicon limitation to the plants. I randomized the location of each pot within the growing space. A flood table bottom watered the pots with nutrient solution. I assigned each plant to one of three herbivory treatments: control, simulated herbivory, or true herbivory. I simulated herbivory by application of 1 mM MeJA solution to the entire above-ground portion of the plant (Waterman et al. 2021b), while crickets housed in water-pik tubes provided true herbivory. Prior to introduction to the the plants, I acclimated crickets by feeding them on the same species used in this trial. Immediately preceding cricket application, I placed them in their tubes and starved them for 24 hours, as this increased the likelihood of the insects initiating feeding rapidly upon exposure to the test plants.

@@

24 hours after treatment application, I harvested leaf material by clipping all undamaged (i.e. no cricket exposure) leaves. I flash froze leaf tissue liquid nitrogen, freeze dried the tissue, and stored it in a -80ºC freezer. To prepare the leaf powder for XRF analysis, I transferred the dried frozen leaf tissue to 2 mL microcentrifuge tube. I cryoground the leaf tissue at -80ºC in a tissuelyser bead mill for 30 seconds at 60 Hz.

@@

To measure the silicon content of the leaf tissue, I followed a modified version of the bench-top XRF method (\cite{reidinger_rapid_2012}). I pressed leaf powder in a hydraulic press at 11 tons of pressure, using a 13mm die to create a pellet. I then placed the pellet in the XRF hood, and used a 30 second scan time to quantify silicon. After each use, I wiped the pellet die and the analyzer clean to minimize contamination between samples.
